---
title: "Integrated Literature Synthesis: Plant Body Mass and Allometric Relationships"
output:
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(readr)
library(dplyr)
library(ggplot2)
library(scales)

safe_read_csv <- function(path) {
  if (file.exists(path)) read_csv(path, show_col_types = FALSE) else tibble::tibble()
}
```

```{r paths}
cwd <- normalizePath(".", winslash = "/", mustWork = TRUE)
if (basename(cwd) == "integrated_literature_synthesis") {
  project_dir <- dirname(cwd)
} else {
  project_dir <- cwd
}
out_dir <- file.path(project_dir, "integrated_literature_synthesis", "output")

papers <- safe_read_csv(file.path(out_dir, "integrated_paper_table.csv"))
study_type <- safe_read_csv(file.path(out_dir, "study_type_summary.csv"))
relationships <- safe_read_csv(file.path(out_dir, "relationship_frequency_summary.csv"))
most_least <- safe_read_csv(file.path(out_dir, "most_least_studied_allometry.csv"))
exp_rel <- safe_read_csv(file.path(out_dir, "exponent_by_relationship_summary.csv"))
evidence_signals <- safe_read_csv(file.path(out_dir, "evidence_signal_summary.csv"))
lang <- safe_read_csv(file.path(out_dir, "language_summary.csv"))
years <- safe_read_csv(file.path(out_dir, "year_summary.csv"))
query_cov <- safe_read_csv(file.path(out_dir, "query_coverage_summary.csv"))
screening <- safe_read_csv(file.path(out_dir, "screening_status_summary.csv"))
stage_counts <- safe_read_csv(file.path(out_dir, "analysis_stage_counts.csv"))

stage_candidate <- "candidate_exploratory"
stage_screened <- "screened_inference_ready"

study_type_candidate <- study_type |> filter(analysis_stage == stage_candidate)
relationships_candidate <- relationships |> filter(analysis_stage == stage_candidate)
most_least_candidate <- most_least |> filter(analysis_stage == stage_candidate)
exp_rel_candidate <- exp_rel |> filter(analysis_stage == stage_candidate)

study_type_screened <- study_type |> filter(analysis_stage == stage_screened)
relationships_screened <- relationships |> filter(analysis_stage == stage_screened)
exp_rel_screened <- exp_rel |> filter(analysis_stage == stage_screened)
```

## Overview

This report integrates automated deep-search outputs into a single quantitative synthesis focused on:

- Study type coverage.
- Allometric relationship coverage.
- Extracted exponent distributions.
- Most and least studied allometry families.

All primary quantitative tables below are shown for candidate-level exploratory records unless explicitly labeled as screened.

```{r overview-metrics}
overview_tbl <- tibble::tibble(
  metric = c(
    "Total integrated papers",
    "Screened included papers",
    "Papers with extracted exponents",
    "Distinct relationship families",
    "Distinct study types"
  ),
  value = c(
    nrow(papers),
    sum(tolower(dplyr::coalesce(papers$include_for_catalog, "")) %in% c("yes", "include", "included"), na.rm = TRUE),
    sum(papers$exponent_count > 0, na.rm = TRUE),
    n_distinct(relationships_candidate$relationship_family),
    n_distinct(study_type_candidate$study_type)
  )
)
knitr::kable(overview_tbl, caption = "Integrated synthesis overview")

if (nrow(stage_counts) > 0) {
  knitr::kable(stage_counts, caption = "Analysis stage counts")
}
```

## Study Type Composition

```{r study-type-table}
study_type_candidate |>
  arrange(desc(paper_n)) |>
  knitr::kable(
    digits = 3,
    caption = "Study type breakdown (candidate exploratory)"
  )
```

```{r study-type-plot, fig.height=4.5}
ggplot(study_type_candidate, aes(x = reorder(study_type, paper_n), y = paper_n)) +
  geom_col(fill = "#2F6B3C") +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(x = NULL, y = "Paper count", title = "Study type coverage") +
  theme_minimal(base_size = 12)
```

## Allometric Relationship Coverage

```{r relationship-top}
relationships_candidate |>
  slice_max(order_by = paper_n, n = 15, with_ties = FALSE) |>
  knitr::kable(caption = "Top 15 most represented relationship families (candidate exploratory)")
```

```{r relationship-bottom}
relationships_candidate |>
  slice_min(order_by = paper_n, n = 10, with_ties = FALSE) |>
  knitr::kable(caption = "10 least represented relationship families (candidate exploratory)")
```

```{r most-least}
most_least_candidate |>
  arrange(rank_class, desc(paper_n)) |>
  knitr::kable(caption = "Most and least studied allometries (candidate exploratory)")
```

## Exponent Synthesis

```{r exponent-overall}
exp_overall <- papers |>
  summarise(
    papers_with_exponents = sum(exponent_count > 0, na.rm = TRUE),
    mean_paper_median_exponent = mean(exponent_median, na.rm = TRUE),
    median_paper_median_exponent = median(exponent_median, na.rm = TRUE),
    min_paper_exponent = min(exponent_min, na.rm = TRUE),
    max_paper_exponent = max(exponent_max, na.rm = TRUE)
  )
knitr::kable(exp_overall, digits = 3, caption = "Overall exponent summary")
```

```{r exponent-by-relationship}
exp_rel_candidate |>
  arrange(desc(n_exponent_values)) |>
  knitr::kable(digits = 3, caption = "Exponent summary by relationship family (candidate exploratory)")
```

```{r evidence-signal-summary}
if (nrow(evidence_signals) > 0) {
  evidence_signals |>
    filter(analysis_stage == stage_candidate) |>
    arrange(desc(paper_n)) |>
    knitr::kable(caption = "Evidence-signal frequencies (not biological relationships)")
}
```

```{r exponent-distribution, fig.height=4.5}
plot_dat <- papers |>
  filter(!is.na(exponent_median))

if (nrow(plot_dat) > 0) {
  ggplot(plot_dat, aes(x = exponent_median)) +
    geom_histogram(binwidth = 0.1, fill = "#6D8A96", color = "white") +
    labs(
      x = "Paper-level median exponent",
      y = "Paper count",
      title = "Distribution of extracted exponents"
    ) +
    theme_minimal(base_size = 12)
}
```

## Language and Time Coverage

```{r language-table}
lang |>
  knitr::kable(digits = 3, caption = "Language composition")
```

## Screened Stage Snapshot

```{r screened-stage}
if (nrow(study_type_screened) > 0 || nrow(relationships_screened) > 0) {
  knitr::kable(study_type_screened, caption = "Study type breakdown (screened inference-ready)")
  knitr::kable(relationships_screened, caption = "Relationship coverage (screened inference-ready)")
  knitr::kable(exp_rel_screened, caption = "Exponent summary (screened inference-ready)")
} else {
  cat("No screened inference-ready records were available in this run.\n")
}
```

```{r publication-year-plot, fig.height=4.5}
if (nrow(years) > 0) {
  ggplot(years, aes(x = publication_year, y = paper_n)) +
    geom_line(color = "#8B4F2E", linewidth = 0.9) +
    geom_point(color = "#8B4F2E", size = 1.6) +
    scale_y_continuous(labels = comma) +
    labs(x = "Publication year", y = "Paper count", title = "Integrated papers over time") +
    theme_minimal(base_size = 12)
}
```

## Retrieval and Screening Diagnostics

```{r query-coverage}
if (nrow(query_cov) > 0) {
  query_cov |>
    arrange(query_id) |>
    knitr::kable(caption = "OpenAlex query coverage diagnostics")
}
```

```{r screening-status}
if (nrow(screening) > 0) {
  screening |>
    knitr::kable(caption = "Current screening status counts")
}
```

## Interpretation Notes

- Candidate-stage relationship and exponent summaries are based on text-mined hits and should be interpreted as screening-level evidence.
- Exponent extraction uses numeric pattern heuristics and may capture values not used as final fitted allometric coefficients.
- Biological relationship tables explicitly exclude method/signal pattern groups.
- Most/least studied designations reflect current mined corpus and query strategy, not the complete global literature.
