---
title: "Deep Literature Search Results Summary"
author: "GitHub Copilot workflow output"
date: "2026-03-30"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(dplyr)
library(readr)
library(knitr)

summary_dir <- normalizePath(dirname(knitr::current_input(dir = TRUE)), winslash = "/", mustWork = TRUE)
output_dir <- file.path(summary_dir, "output")
screening_dir <- file.path(summary_dir, "screening")

run_parameters <- read_csv(file.path(output_dir, "run_parameters.csv"), show_col_types = FALSE)
query_log <- read_csv(file.path(output_dir, "search_query_log.csv"), show_col_types = FALSE)
ranked <- read_csv(file.path(output_dir, "openalex_deep_ranked_candidates.csv"), show_col_types = FALSE)
pattern_hits <- read_csv(file.path(output_dir, "openalex_pattern_hits_long.csv"), show_col_types = FALSE)
screening_summary <- read_csv(file.path(screening_dir, "deep_search_screening_queue_summary.csv"), show_col_types = FALSE)
screening_queue <- read_csv(file.path(screening_dir, "deep_search_screening_queue.csv"), show_col_types = FALSE)

run_row <- run_parameters[1, ]

top_journals <- ranked |>
  count(journal, sort = TRUE, name = "ranked_papers") |>
  mutate(journal = ifelse(is.na(journal) | journal == "", "Unknown", journal)) |>
  slice_head(n = 15)

top_patterns <- pattern_hits |>
  count(pattern_group, sort = TRUE, name = "hit_rows") |>
  slice_head(n = 15)

queries_needing_expansion <- query_log |>
  filter(terminal_reason == "max_pages_reached") |>
  select(query_id, query_text, records_fetched, terminal_reason)

top_candidates <- ranked |>
  mutate(
    doi_link = ifelse(
      !is.na(doi) & nzchar(doi),
      sprintf("<a href=\"%s\">doi</a>", doi),
      ""
    )
  ) |>
  select(
    title,
    publication_year,
    journal,
    cited_by_count,
    n_pattern_groups,
    hit_groups,
    deep_relevance_score,
    doi_link
  ) |>
  slice_head(n = 25)

screening_preview <- screening_queue |>
  select(
    screening_rank,
    screening_priority,
    full_text_priority,
    title,
    publication_year,
    journal,
    include_for_catalog,
    accepted_status
  ) |>
  slice_head(n = 25)
```

## Overview

This report summarizes the first hardened deep-search run for plant body-mass allometry literature.

- Run UTC: `r run_row$run_utc`
- Script version: `r run_row$script_version`
- Query families: `r run_row$n_query_families`
- Pattern groups: `r run_row$n_pattern_groups`
- Precision rule: `r run_row$precision_rule`

## Headline Counts

```{r headline-counts}
headline_tbl <- tibble::tibble(
  Metric = c(
    "Raw candidates fetched",
    "Pattern-hit rows",
    "Ranked unique papers before gate",
    "Ranked unique papers after gate",
    "High-priority screening papers",
    "Priority full-text papers"
  ),
  Value = c(
    run_row$raw_candidate_n,
    run_row$pattern_hit_rows_n,
    run_row$ranked_pre_gate_n,
    run_row$ranked_post_gate_n,
    screening_summary$high_priority_n,
    screening_summary$priority_full_text_n
  )
)

kable(headline_tbl, format = "html", caption = "Main output counts from the deep-search run.")
```

## Query Performance

```{r query-performance}
query_tbl <- query_log |>
  select(
    query_id,
    query_text,
    pages_attempted,
    pages_ok,
    pages_failed,
    records_fetched,
    terminal_reason
  )

kable(query_tbl, format = "html", caption = "Per-query retrieval summary.")
```

## Queries That Hit the Page Cap

```{r query-cap}
if (nrow(queries_needing_expansion) == 0) {
  cat("No queries hit the configured page cap in this run.")
} else {
  kable(
    queries_needing_expansion,
    format = "html",
    caption = "Queries that hit max_pages_reached and may merit deeper expansion in a future run."
  )
}
```

## Top Journals In Ranked Results

```{r top-journals}
kable(top_journals, format = "html", caption = "Journals most represented in the post-gate ranked candidate set.")
```

## Most Frequent Pattern Groups

```{r top-patterns}
kable(top_patterns, format = "html", caption = "Pattern groups most frequently matched across abstract-hit rows.")
```

## Top Ranked Candidate Papers

```{r top-candidates}
kable(
  top_candidates,
  format = "html",
  escape = FALSE,
  digits = 2,
  caption = "Top 25 ranked candidate papers from the post-gate set."
)
```

## Screening Queue Summary

```{r screening-summary}
kable(screening_summary, format = "html", caption = "Summary of the generated screening queue.")
```

## Screening Queue Preview

```{r screening-preview}
kable(screening_preview, format = "html", caption = "First 25 rows from the manual screening queue.")
```

## Interpretation Notes

- The deep-search workflow found a large candidate set and retained `r run_row$ranked_post_gate_n` papers after the precision gate.
- Several broad query families hit `max_pages_reached`, so this is a substantial first pass rather than a guaranteed exhaustive crawl.
- The screening queue is now ready for manual triage, taxonomic reconciliation, and integration into the main catalog workflow.