---
title: "Plant Body-Mass Scaling Relationships: Comprehensive Working Catalog"
author: "Working synthesis"
date: "2026-03-28"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

## Purpose

This is a comprehensive working catalog of measured plant allometric relationships linking traits or life-history attributes to plant size (mass, diameter, height, or related size proxies) across physiology, ecology, and evolution.

## Important Scope Note

No synthesis can guarantee inclusion of literally every estimate ever published. This document is designed as a high-coverage, citable starting point and should be iteratively expanded as you encounter clade-, biome-, and method-specific studies.

## Systematic Search Workflow (Updated)

This workflow is designed to move this catalog toward near-exhaustive, reproducible coverage.

### Databases and Sources

- Web of Science Core Collection
- Scopus
- BIOSIS
- CAB Abstracts
- AGRICOLA
- JSTOR
- Google Scholar (first 300 relevance-ranked records per query)
- Crossref/OpenAlex (for DOI completion and citation graph expansion)
- Forward and backward citation chaining from core syntheses

### Search Strings

```{r}
search_queries <- tibble::tribble(
  ~Query_id, ~Focus, ~Search_string,
  "A", "General allometry", "(plant OR tree OR shrub OR herb OR angiosperm OR gymnosperm OR fern OR bryophyte) AND (allometr* OR \"scaling relationship*\" OR \"power law\" OR exponent) AND (biomass OR mass OR \"stem diameter\" OR DBH OR height)",
  "B", "Physiology", "(plant*) AND (metabolic rate OR respiration OR photosynthesis OR transpiration OR hydraulic conductance) AND (allometr* OR exponent OR scaling) AND (biomass OR diameter OR height)",
  "C", "Ecology and demography", "(plant*) AND (\"self-thinning\" OR mortality OR recruitment OR fecundity OR \"seed output\" OR abundance) AND (allometr* OR size-scaling OR exponent)",
  "D", "Evolution and trait integration", "(plant*) AND (\"seed mass\" OR \"leaf size\" OR twig OR fruit mass OR reproductive allocation) AND (allometr* OR \"size scaling\") AND (phylogen* OR comparative)",
  "E", "Methods and model fitting", "(SMA OR RMA OR \"standardized major axis\" OR \"mixed effects\" OR \"phylogenetic regression\") AND (plant allometr*)"
)

knitr::kable(search_queries, format = "html", escape = TRUE, caption = "Recommended search queries for updating this catalog.")
```

### Screening and Eligibility Rules

Include studies that:

- Report empirical plant scaling fits or provide enough data to recover slope/exponent.
- Provide trait definition, size metric, taxonomic scope, and analyzable units.
- Identify sampling context (clade, biome, ontogenetic stage, or experiment type).

Exclude studies that:

- Are non-plant only or do not provide empirical plant fit data.
- Are purely theoretical without plant calibration.
- Lack metadata needed to map relationship type, scope, or units.

### Deep-Search Augmentation Handoff

When using the abstract-mining augmentation pipeline (`deep_literature_search/`), apply this handoff before adding rows to the catalog:

1. Prioritize papers from `openalex_deep_ranked_candidates.csv`.
2. Confirm each candidate is truly plant-focused and contains empirical allometric evidence (fitted relationship or recoverable exponent).
3. Record why any high-ranked candidate was excluded (for reproducibility and precision auditing).
4. Attach taxonomy reconciliation fields before integration:
  - reported taxon name
  - accepted taxon name
  - accepted status (accepted/synonym/unresolved)
  - backbone source + version date
  - backbone ID, match method, and confidence
5. Map accepted studies into the same extraction schema used below so curated and deep-search streams remain comparable.

### Taxonomy Reconciliation Requirements

- Use an explicit plant backbone (for example POWO/WFO), with version/date.
- Preserve original reported taxon names and store reconciled accepted names separately.
- Record accepted status, synonym status, backbone ID, match method, confidence, and timestamp.
- Keep unresolved or ambiguous names explicitly; never force-match silently.

### Evidence Tiering

Each relationship should be tagged into one tier:

- Empirical quantitative
- Empirical qualitative
- Conceptual or historical

This reduces the risk of treating conceptual statements as equivalent to fitted quantitative evidence.

## Column Definitions

- `Typical_scaling_form`: Common mathematical form used in the literature.
- `First_discovery_year`: Earliest commonly cited seminal source for that relationship in plant biology.
- `Key_citations`: Seminal or widely used references.
- `Explanation`: Plain-language interpretation.
- `Core_assumptions`: Assumptions required for meaningful inference.

## Comprehensive Table

```{r}
library(tibble)
library(dplyr)
library(knitr)

scaling_tbl <- tribble(
  ~Domain, ~Relationship, ~Typical_scaling_form, ~First_discovery_year, ~Key_citations, ~Explanation, ~Core_assumptions,

  "Physiology",
  "Whole-plant metabolic rate vs biomass",
  "B = a*M^b, b ~ 0.75",
  1998,
  "Enquist et al. (1998); West et al. (1999)",
  "Larger plants use more total energy but sublinearly per unit mass.",
  "Transport network is approximately space-filling; terminal units and resource delivery constraints are comparable.",

  "Physiology",
  "Whole-plant respiration vs biomass",
  "R = a*M^b, b ~ 0.75-1.0",
  2006,
  "Reich et al. (2006); Mori et al. (2010)",
  "Respiration increases with size, with exponent shifts by tissue and size class.",
  "Temperature, tissue composition, and measurement protocol are normalized or modeled.",

  "Physiology",
  "Whole-plant photosynthesis vs biomass",
  "P = a*M^b, b ~ 0.7-1.0",
  1998,
  "Enquist et al. (1998); Niklas and Enquist (2001)",
  "Carbon gain rises with size, often less than proportionally.",
  "Comparable light environment and canopy structure, or explicit control for them.",

  "Physiology",
  "Absolute growth rate vs biomass",
  "AGR = a*M^b, b ~ 0.75",
  2001,
  "Niklas and Enquist (2001); Enquist et al. (2007)",
  "Larger plants add more mass per unit time in absolute terms.",
  "Growth near quasi-steady resource allocation and comparable time windows.",

  "Physiology",
  "Relative growth rate vs biomass",
  "RGR = a*M^b, b ~ -0.25",
  2001,
  "Niklas and Enquist (2001)",
  "Mass-specific growth usually declines with increasing size.",
  "Ontogenetic effects are controlled or explicitly modeled.",

  "Physiology",
  "Time to maturation vs biomass",
  "T = a*M^b, b ~ 0.25",
  2001,
  "Niklas and Enquist (2001); West et al. (1999)",
  "Larger-bodied plants generally take longer to reproduce.",
  "Comparable life-history stage definitions and reproductive onset criteria.",

  "Physiology",
  "Lifespan vs biomass",
  "L = a*M^b, b ~ 0.15-0.30",
  2001,
  "Niklas and Enquist (2001)",
  "Longevity tends to increase with maximum plant size.",
  "Mortality hazards and disturbance regimes are comparable or modeled.",

  "Physiology",
  "Sapwood area vs supported leaf area (pipe model)",
  "A_s = a*L_A^b, b ~ 1",
  1964,
  "Shinozaki et al. (1964a, 1964b)",
  "Conductive stem area tracks distal leaf demand.",
  "Hydraulic supply-demand matching and valid sapwood identification.",

  "Physiology",
  "Whole-plant transpiration vs sapwood area",
  "E = a*A_s^b, b ~ 1",
  1984,
  "Waring et al. (1984); Meinzer et al. (2005)",
  "Water flux scales with conducting cross-sectional area.",
  "VPD and stomatal regulation differences are controlled or represented.",

  "Physiology",
  "Stem diameter vs aboveground biomass",
  "AGB = a*D^b, b ~ 2-3",
  1994,
  "Niklas (1994); Chave et al. (2014)",
  "Diameter is a strong predictor of biomass.",
  "Wood density and height effects are included or constrained.",

  "Physiology",
  "Height vs stem diameter",
  "H = a*D^b, b ~ 0.5-0.8",
  1976,
  "McMahon and Kronauer (1976); Niklas (1994)",
  "Taller plants require disproportionate stem support and hydraulic capacity.",
  "Mechanical stability and hydraulic constraints jointly shape observed scaling.",

  "Physiology",
  "Crown radius vs stem diameter",
  "C_r = a*D^b, b ~ 0.5-1.0",
  1990,
  "King (1990); Niklas (1994)",
  "Crown spread scales predictably with stem size.",
  "Limited distortion from extreme asymmetric competition.",

  "Physiology",
  "Crown volume vs stem diameter",
  "V_c = a*D^b, b ~ 1.5-3",
  1990,
  "King (1990); West et al. (2009)",
  "Canopy volume increases steeply with trunk size.",
  "Crown architecture class is consistent enough for pooled inference.",

  "Physiology",
  "Leaf mass vs stem mass",
  "M_L = a*M_S^b, b ~ 0.7-1.1",
  2002,
  "Enquist and Niklas (2002); Poorter et al. (2012)",
  "Biomass partitioning between leaves and stems shifts predictably with size.",
  "Comparable ontogenetic stage and allocation context.",

  "Physiology",
  "Root mass vs shoot mass",
  "M_R = a*M_Sh^b, b ~ 0.8-1.2",
  2002,
  "Enquist and Niklas (2002); Poorter et al. (2012)",
  "Belowground and aboveground investment covary allometrically.",
  "Soil resource heterogeneity is accounted for or limited.",

  "Physiology",
  "Fine-root respiration vs root biomass",
  "R_r = a*M_r^b, b ~ 0.75-1",
  2006,
  "Reich et al. (2006)",
  "Root maintenance costs rise with root system size.",
  "Comparable root turnover and tissue activity classes.",

  "Physiology",
  "Vessel diameter vs stem diameter",
  "d_v = a*D^b, b > 0",
  2006,
  "Anfodillo et al. (2006); Olson et al. (2014)",
  "Larger stems tend to contain wider conduits.",
  "Hydraulic safety-efficiency tradeoff is the dominant design axis.",

  "Physiology",
  "Conduit tapering vs path length or height",
  "d_v = a*L^b with axial taper",
  1999,
  "West et al. (1999); Savage et al. (2010)",
  "Conduit tapering can offset path-length resistance.",
  "Network architecture approximates transport-theory assumptions.",

  "Physiology",
  "Bark thickness vs stem diameter",
  "T_b = a*D^b, b ~ 0.8-1.2",
  1983,
  "Jackson et al. (1999); Rosell et al. (2017)",
  "Protective bark investment scales with stem size and disturbance pressure.",
  "Disturbance regime context (fire/herbivory) is represented.",

  "Physiology",
  "Petiole diameter vs lamina size",
  "D_p = a*A_leaf^b, b ~ 0.4-0.7",
  1949,
  "Corner (1949); Olson et al. (2009)",
  "Petiole support and hydraulic capacity track lamina dimensions.",
  "Leaf mechanical design and venation constraints are comparable.",

  "Ecology",
  "Self-thinning: stand density vs mean plant mass",
  "N = a*Mbar^(-3/2) (classic)",
  1963,
  "Yoda et al. (1963); Weller (1987)",
  "Crowded stands lose individuals as mean size increases.",
  "Competition is strong and acts on shared limiting resources.",

  "Ecology",
  "Stand density vs mean stem diameter",
  "N = a*D^b, b ~ -1.5 to -2",
  1933,
  "Reineke (1933); Weller (1987)",
  "Density-size tradeoff represented with diameter metric.",
  "Even-aged stand approximation and site comparability.",

  "Ecology",
  "Individual light interception vs plant size",
  "I = a*M^b, often b < 1",
  1990,
  "King (1990); Falster and Westoby (2003)",
  "Larger plants intercept more light but often with diminishing returns.",
  "Neighborhood and crown-geometry effects are modeled or constrained.",

  "Ecology",
  "Mortality rate vs stem diameter",
  "mu = a*D^b, often b < 0",
  1995,
  "Kobe et al. (1995); Lines et al. (2010)",
  "Background mortality often declines with size.",
  "Acute disturbance mortality is separated from background mortality.",

  "Ecology",
  "Recruitment output vs parent size",
  "F = a*M^b, b > 0",
  2006,
  "Muller-Landau et al. (2006); Moles et al. (2008)",
  "Larger plants tend to produce more propagules and recruits.",
  "Seed limitation and microsite limitation are distinguished.",

  "Ecology",
  "Seed output vs plant biomass",
  "N_seed = a*M^b, b ~ 0.7-1.2",
  1992,
  "Aarssen and Taylor (1992); Henery and Westoby (2001)",
  "Reproductive output increases strongly with plant size.",
  "Comparable reproductive season and pollination context.",

  "Ecology",
  "Forest individual production vs size",
  "P_i = a*M^b, b ~ 0.75",
  2007,
  "Enquist et al. (2007)",
  "Large individuals contribute disproportionately to stand production.",
  "Census intervals and growth estimation are comparable.",

  "Ecology",
  "Gap-capture ability vs size",
  "G = a*D^b, b > 0",
  2005,
  "Poorter et al. (2005); Wright et al. (2010)",
  "Larger crowns and stems improve post-disturbance light capture.",
  "Disturbance regime and local competitive structure are represented.",

  "Ecology",
  "Species abundance vs typical adult size",
  "A = a*M^b, usually b < 0",
  2006,
  "Muller-Landau et al. (2006); White et al. (2007)",
  "Communities often contain fewer individuals of larger species.",
  "Sampling completeness and detectability bias are addressed.",

  "Ecology",
  "Growth-mortality tradeoff with size",
  "g(D) and mu(D) follow size-dependent allometries",
  2011,
  "Ruger et al. (2011); Iida et al. (2014)",
  "Size mediates demographic tradeoffs among ecological strategies.",
  "Reliable repeated censuses and sufficient demographic events.",

  "Evolution",
  "Seed mass vs adult plant height",
  "M_seed = a*H^b, b > 0",
  2005,
  "Moles et al. (2005); Westoby et al. (1992)",
  "Taller species tend to evolve larger seeds on average.",
  "Phylogenetic non-independence is explicitly modeled.",

  "Evolution",
  "Seed number-seed mass tradeoff",
  "N_seed proportional to M_seed^(-1) at fixed reproductive mass",
  1974,
  "Smith and Fretwell (1974); Moles and Westoby (2006)",
  "Larger offspring generally reduce offspring number under budget constraints.",
  "Approximate fixed reproductive allocation per episode.",

  "Evolution",
  "Leaf size vs twig size (Corner's rule)",
  "A_leaf = a*D_twig^b, b > 0",
  1949,
  "Corner (1949); Sun et al. (2006)",
  "Evolutionary integration links support axes and leaf dimensions.",
  "Developmental integration constraints are comparable across taxa.",

  "Evolution",
  "Branching area preservation across sizes",
  "Sum daughter diameters^p approximately parent diameter^p",
  1500,
  "Leonardo da Vinci notes; Eloy (2011)",
  "Branching allometry recurs across lineages and scales.",
  "Hydraulic or mechanical optimality dominates branching design.",

  "Evolution",
  "Maximum height vs stem diameter across clades",
  "H_max = a*D^b, b < 1",
  1994,
  "Niklas (1994); Niklas and Spatz (2004)",
  "Mechanical and hydraulic constraints bound height evolution.",
  "Environment and phylogeny are controlled in comparative analyses.",

  "Evolution",
  "Reproductive biomass vs vegetative size",
  "M_repr = a*M^b, b context dependent",
  1998,
  "Sugiyama and Bazzaz (1998); Weiner (2004)",
  "Allocation schedules evolve with life history and size trajectory.",
  "Ontogenetic stage and resource context are standardized.",

  "Evolution",
  "Xylem conduit size vs plant size across angiosperms",
  "d_v = a*D^b, b > 0",
  2009,
  "Olson et al. (2009); Olson and Rosell (2013)",
  "Conduit scaling reflects adaptation to size-driven transport demand.",
  "Climate and habitat effects are partitioned from pure size effects.",

  "Evolution",
  "Floral display size vs vegetative size",
  "F_size = a*M^b, usually shallow positive b",
  1997,
  "Delph et al. (1997); Sletvold (2002)",
  "Larger plants generally support larger or more numerous floral displays.",
  "Pollination environment and mating system differences are modeled.",

  "Evolution",
  "Fruit mass vs plant size",
  "M_fruit = a*M^b, b > 0",
  2005,
  "Bolmgren and Eriksson (2005); Moles et al. (2007)",
  "Fruit size often increases with overall plant size and dispersal syndrome.",
  "Phylogeny and disperser-guild effects are represented.",

  "Cross-domain",
  "Leaf lifespan vs LMA and whole-plant size strategy",
  "LL = a*LMA^b (positive), with size-strategy covariation",
  1992,
  "Reich et al. (1992); Wright et al. (2004)",
  "Leaf economics relationships integrate with size-dependent growth strategies.",
  "Trait protocols are standardized and canopy position effects are controlled.",

  "Cross-domain",
  "Hydraulic conductance vs stem size",
  "K = a*D^b",
  1985,
  "Tyree and Zimmermann (1985) and later syntheses",
  "Larger stems generally have disproportionately larger conductance.",
  "Temperature, viscosity, path length, and embolism status are accounted for.",

  "Cross-domain",
  "Whole-plant mass vs stem diameter",
  "M = a*D^b",
  1994,
  "Niklas (1994); Niklas and Spatz (2004)",
  "Diameter-based scaling underpins many trait-size and demographic allometries.",
  "Allometric model form and wood-density variation are appropriately handled."
)

kable(
  scaling_tbl,
  format = "html",
  escape = TRUE,
  caption = "Comprehensive working catalog of measured plant size-scaling relationships."
)
```

## Second Table: Confidence, Exponent Ranges, and Analysis Use

```{r}
confidence_tbl <- tribble(
  ~Relationship, ~Evidence_confidence, ~Typical_exponent_range, ~Typical_data_scope, ~Analysis_use_notes,

  "Whole-plant metabolic rate vs biomass", "High", "b ~ 0.70-0.85", "Cross-species and ontogenetic datasets", "Strong for broad comparative scaling; avoid assuming a universal fixed exponent in every clade.",
  "Whole-plant respiration vs biomass", "High", "b ~ 0.75-1.00", "Seedlings to mature plants; tissue-specific contrasts", "Report temperature correction and tissue composition assumptions.",
  "Whole-plant photosynthesis vs biomass", "Moderate", "b ~ 0.70-1.00", "Cross-species and environment-stratified datasets", "Include light environment and canopy architecture covariates.",
  "Absolute growth rate vs biomass", "High", "b ~ 0.65-0.85", "Ontogenetic growth series; multispecies compilations", "Use consistent growth intervals and distinguish stress vs non-stress regimes.",
  "Relative growth rate vs biomass", "High", "b ~ -0.35 to -0.15", "Seedling and juvenile datasets; ontogenetic gradients", "Control for ontogeny explicitly when comparing taxa.",
  "Sapwood area vs supported leaf area", "High", "b ~ 0.9-1.1", "Tree branch/segment and whole-tree measurements", "Useful hydraulic proxy; validate sapwood-heartwood separation protocol.",
  "Stem diameter vs aboveground biomass", "High", "b ~ 2.0-3.2", "Forest inventories, destructive calibration studies", "Core equation for biomass estimation; include wood density and height when possible.",
  "Height vs stem diameter", "High", "b ~ 0.5-0.9", "Forest plots and species-level allometry datasets", "Interpret as joint hydraulic-mechanical outcome; stratify by life form.",
  "Root mass vs shoot mass", "High", "b ~ 0.8-1.2", "Pot and field studies across resource gradients", "Report soil/resource context because slope and intercept can shift.",
  "Vessel diameter vs stem diameter", "Moderate", "b > 0 (often ~0.1-0.4)", "Xylem anatomical datasets across stem sizes", "Pair with embolism/safety traits for mechanistic interpretation.",
  "Self-thinning: stand density vs mean mass", "High", "slope ~ -1.5 (classical, variable)", "Even-aged dense stands under competition", "Use cautiously outside strong crowding or mixed-age stands.",
  "Stand density vs mean diameter", "High", "b ~ -1.5 to -2.0", "Forestry and long-term stand datasets", "Best interpreted near upper-density frontier conditions.",
  "Mortality rate vs stem diameter", "Moderate", "b often < 0", "Repeated forest censuses", "Separate background from disturbance mortality for inference.",
  "Seed output vs plant biomass", "Moderate", "b ~ 0.7-1.2", "Reproductive census datasets", "Estimate over complete reproductive windows; account for masting where relevant.",
  "Species abundance vs adult size", "Moderate", "b typically < 0", "Community inventory datasets", "Correct for detectability and sampling completeness before cross-site comparisons.",
  "Seed mass vs adult height", "High", "b > 0", "Global trait compilations with phylogeny", "Apply phylogenetic models; raw regressions can overstate significance.",
  "Seed number-seed mass tradeoff", "High", "near -1 at fixed reproductive budget", "Comparative reproductive allocation datasets", "Interpret within explicit reproductive budget assumptions.",
  "Leaf size vs twig size (Corner's rule)", "High", "b > 0", "Comparative morphology datasets", "Strong structural integration signal; useful in trait-network analyses.",
  "Maximum height vs stem diameter", "High", "b < 1", "Broad clade and biome datasets", "Interpret with climatic and biome context to avoid overgeneralization.",
  "Reproductive biomass vs vegetative size", "Moderate", "b context dependent", "Life-history and allocation datasets", "Separate annual vs perennial strategies and ontogenetic stage.",
  "Fruit mass vs plant size", "Moderate", "b > 0", "Comparative reproductive trait datasets", "Model phylogeny and dispersal syndrome jointly for robust inference.",
  "Hydraulic conductance vs stem size", "High", "positive scaling; b variable", "Stem hydraulic measurements across sizes", "Standardize for temperature, segment length, and embolism state.",
  "Whole-plant mass vs stem diameter", "High", "b commonly > 2", "Destructive biomass and inventory datasets", "Foundational bridge trait for translating diameter to multiple performance variables."
)

kable(
  confidence_tbl,
  format = "html",
  escape = TRUE,
  caption = "Evidence confidence, exponent ranges, and practical analysis notes for key scaling relationships."
)
```

## Third Table: Study-Level Extraction Schema (For Systematic Review)

```{r}
extraction_schema_tbl <- tibble::tribble(
  ~Field_name, ~Required, ~Description, ~Example,
  "record_id", "Yes", "Unique row ID for each extracted relationship", "PBMSR_0001",
  "source_type", "Yes", "Primary empirical, meta-analysis, theory, or review", "Primary empirical",
  "citation_short", "Yes", "Short reference used in tables", "Niklas and Enquist (2001)",
  "doi", "Recommended", "DOI for provenance and deduplication", "10.1038/35035055",
  "year", "Yes", "Publication year", "2001",
  "relationship_name", "Yes", "Human-readable name of scaling relationship", "Absolute growth rate vs biomass",
  "domain", "Yes", "Physiology, ecology, evolution, or cross-domain", "Physiology",
  "response_trait", "Yes", "Trait on y-axis of scaling model", "Absolute growth rate",
  "size_predictor", "Yes", "Size variable on x-axis", "Whole-plant dry mass",
  "scaling_form", "Yes", "Model form exactly as fit or reported", "AGR = a*M^b",
  "exponent_b", "Recommended", "Estimated scaling exponent", "0.74",
  "exponent_ci", "Recommended", "Confidence interval (or SE) for exponent", "0.70-0.78",
  "intercept_a", "Optional", "Intercept estimate in reported units", "0.12",
  "fit_method", "Yes", "OLS, SMA, RMA, mixed-effects, phylogenetic model", "SMA",
  "sample_size_n", "Yes", "Number of observations used in fit", "312",
  "taxon_scope_reported", "Yes", "Original taxonomic scope as stated", "Angiosperms",
  "name_input_verbatim", "Recommended", "Original taxon names exactly as reported", "Quercus agrifolia",
  "name_accepted", "Recommended", "Reconciled accepted name", "Quercus agrifolia Née",
  "taxon_id", "Recommended", "Backbone taxon ID", "POWO:urn:lsid:ipni.org:names:296583-1",
  "taxonomic_status", "Recommended", "Accepted, synonym, unresolved, ambiguous", "Accepted",
  "backbone_source", "Recommended", "Taxonomic backbone used for reconciliation", "POWO",
  "backbone_version_date", "Recommended", "Backbone version or access date", "2026-03-29",
  "biome_or_system", "Recommended", "Biome/region/system context", "Temperate forest",
  "ontogenetic_stage", "Recommended", "Seedling, sapling, mature, mixed", "Mixed",
  "trait_units", "Yes", "Units of response trait", "g day^-1",
  "size_units", "Yes", "Units of size predictor", "g",
  "temperature_or_env_controls", "Optional", "Key environmental normalization variables", "Temperature normalized to 25C",
  "notes_assumptions", "Yes", "Study-specific assumptions needed for interpretation", "Comparable ontogenetic stages",
  "evidence_tier", "Yes", "Empirical quantitative, empirical qualitative, conceptual/historical", "Empirical quantitative",
  "quality_flag", "Recommended", "High, moderate, low extraction confidence", "High",
  "data_source_url", "Yes", "URL to publication or data source", "https://doi.org/10.1038/35035055",
  "extracted_by", "Yes", "Person/agent who extracted row", "GitHub Copilot",
  "extracted_utc", "Yes", "Extraction timestamp in UTC", "2026-03-29 15:25:00 UTC"
)

knitr::kable(
  extraction_schema_tbl,
  format = "html",
  escape = TRUE,
  caption = "Recommended study-level extraction schema for reproducible scaling synthesis."
)
```

## Interpretation Notes

- Exponents are often context-dependent and vary by clade, biome, ontogenetic stage, and statistical method.
- `First_discovery_year` indicates the commonly cited seminal point for each relationship, not necessarily the earliest historical mention.

## Suggested Next Expansion

1. Add clade-specific rows (angiosperms, gymnosperms, ferns, bryophytes, crops).
2. Add biome-specific exponent tables and confidence intervals.
3. Add method metadata (OLS vs SMA/RMA vs mixed-effects vs phylogenetic models).
4. Add uncertainty fields per relationship (estimate range, sample size, taxa count).
